\Large \bf Uncovering Bugs in Distributed Storage Systems\\during Testing (not in Production!)

### 2015-12-22 evaluation.tex
update

\setlength{\tabcolsep}{.72em}
\begin{table*}[t]
\small
\centering
\begin{tabular}{rl rrr rrr}
\centering
\input{experiments/tables/results.tex}
\end{tabular}
\caption{Results from running the \psharp random and PCT systematic testing schedulers for 100,000 iterations. We report: time in seconds to find a bug (Time to Bug); number of scheduling steps when a bug was found (\#SS); and if a bug was found with a particular scheduler (BF?).}
\label{tab:testing}
\end{table*}

@@

We performed all experiments using the Windows PowerShell tool on a 2.50GHz Intel Core i5-4300U CPU with 8GB RAM running Windows 10 Pro 64-bit. We configured the engine to perform 100,000 iterations. The random seed for both schedulers was generated using the current time. The PCT scheduler was further configured with a bug depth of 2 and a max number of scheduling steps of 500. All reported times are in seconds.

@@

For the MigratingTable case study, we evaluated the \psharp test harness of \S\ref{sec:cases:migration} on eleven buggy versions of the system, including eight \emph{organic} bugs that actually occurred in development and three \emph{notional} bugs (denoted by $*$), which are other code changes that we deemed interesting ways of making the system incorrect. The test found seven of the organic bugs, which are listed first in Table~\ref{tab:testing}. The remaining four bugs (denoted by $\diamond$) were not caught with our default test harness in the 100,000 iterations (a process that required less than 30 minutes). We believe this is because the inputs and schedules that trigger them are too rare in our distribution. To confirm this, we wrote a custom test case for each bug with a specific input that triggers it and were able to quickly reproduce the bugs, as shown. This method is useful for regression testing, but one could not use it during development with no prior knowledge about the bugs.

>>>>

\setlength{\tabcolsep}{.72em}
\begin{table*}[t]
\small
\centering
\begin{tabular}{rl rrr rrr}
\centering
\input{experiments/tables/results.tex}
\end{tabular}
\caption{Results from running the \psharp random and PCT systematic testing schedulers for 100,000 executions. We report: time in seconds to find a bug (Time to Bug); number of scheduling steps when a bug was found (\#SS); and if a bug was found with a particular scheduler (BF?).}
\label{tab:testing}
\end{table*}

@@

We performed all experiments using the Windows PowerShell tool on a 2.50GHz Intel Core i5-4300U CPU with 8GB RAM running Windows 10 Pro 64-bit. We configured the engine to perform 100,000 executions. The random seed for both schedulers was generated using the current time. The PCT scheduler was further configured with a bug depth of 2 and a max number of scheduling steps of 500. All reported times are in seconds.

@@

For the MigratingTable case study, we evaluated the \psharp test harness of \S\ref{sec:cases:migration} on eleven buggy versions of the system, including eight \emph{organic} bugs that actually occurred in development and three \emph{notional} bugs (denoted by $*$), which are other code changes that we deemed interesting ways of making the system incorrect. The test found seven of the organic bugs, which are listed first in Table~\ref{tab:testing}. The remaining four bugs (denoted by $\diamond$) were not caught with our default test harness in the 100,000 executions (a process that required less than 30 minutes). We believe this is because the inputs and schedules that trigger them are too rare in our distribution. To confirm this, we wrote a custom test case for each bug with a specific input that triggers it and were able to quickly reproduce the bugs, as shown. This method is useful for regression testing, but one could not use it during development with no prior knowledge about the bugs.

### 2015-12-22 methodology.tex
update

The original \psharp trace did not include sufficient details to allow the developers to identify the root cause of the problem. Fortunately, running the test harness took very little time, so the developers were able to quickly iterate and add more refined debuggin outputs in each iteration. After several iterations, the developers were able to pinpoint the exact culprit and immediately propose a solution for fixing the bug. Once the proposed solution was implemented, the developers ran the test harness again. No bugs were found for 100,000 executions, a process that only took on the order of tens of minutes.

>>>>

The original \psharp trace did not include sufficient details to allow the developers to identify the root cause of the problem. Fortunately, running the test harness took very little time, so the developers were able to quickly iterate and add more refined debugging outputs in each iteration. After several iterations, the developers were able to pinpoint the exact culprit and immediately propose a solution for fixing the bug. Once the proposed solution was implemented, the developers ran the test harness again. No bugs were found for 100,000 executions, a process that only took on the order of tens of minutes.

### 2015-12-22 motivation.tex
update

\noindent
Despite of the extensive testing efforts, the vNext developers have been plagued for months by an elusive bug in the ExtMgr logic. All the unit test suites and integration test suites successfully pass every single time. However, the stress test suite fails \emph{from time to time} after very long executions; in these cases, certain replicas of some extents fail and are never repaired. This bug has proven to be difficult to identify, reproduce and troubleshoot. First, it requires very long executions to trigger. Second, an extent not being repaired is \emph{not} a property that can be easily verified. In practice, the developers rely on a very large time-out period to detect the bug. Finally, by the time that the bug is detected, very long execution traces have been collected, which makes manual inspection tedious and ineffective.

To uncover this bug, as well as many other tricky bugs, the vNext developers are in constant search of a generic and systematic approach for testing production-scale distributed storage systems.

%Since this is a rather typical dilemma in the development of distributed storage systems, a general and systematic approach hopefully would be very helpful to the developers and greatly increase their productivity.

>>>>

\noindent
Despite of the extensive testing efforts, the vNext developers have been plagued for months by an elusive bug in the ExtMgr logic. All the unit test suites and integration test suites successfully pass every single time. However, the stress test suite fails \emph{from time to time} after very long executions; in these cases, certain replicas of some extents fail and are never repaired.

This bug has proven to be difficult to identify, reproduce and troubleshoot. First, an extent never being repaired is \emph{not} a property that can be easily checked. Second, the bug seems to manifest only in very long executions. Finally, by the time that the bug is detected, very long execution traces have been collected, which makes manual inspection tedious and ineffective. To uncover this bug, as well as many other tricky bugs, the vNext developers are in constant search of a generic and systematic approach for testing distributed systems.

### 2015-12-22 overview.tex
update

The goal of our work is to develop testing techniques for detecting and debugging Heisenbugs in distributed storage systems prior to deployment. To achieve our goal, we use \psharp~\cite{deligiannis2015psharp}, an extension of the mainstream \csharp language that provides: (i) language support for \emph{modeling the environment} of distributed systems developed using Microsoft's .NET framework; and (ii) a \emph{testing engine} that can systematically explore all interleavings between asynchronous event handlers, as well as other nondeterministic events such as failures and timeouts.

@@

\item The environment of the system must be modeled in \psharp, which offers APIs for obtaining non-deterministically chosen values. These values can be used to simulate failures, timers, etc. (see \S\ref{sec:method:timer} and \S\ref{sec:method:driver}). This step captures external sources of nondeterminism. The \psharp testing engine can then systematically explore all choices, enumerating different asynchronous event interleavings, possible points of failures, timers firing, etc. (see \S\ref{sec:psharp:testing}).

\item The criteria for correctness of an execution must be specified. Specifications in \psharp can encode either \emph{safety} or \emph{liveness}~\cite{lamport1977proving} properties (see \S\ref{sec:bg:bugs}). Safety specifications generalize the notion of source code assertions; a safety violation is a finite trace leading to an erroneous state. Liveness specifications generalize nontermination; a liveness violation is an infinite trace that exhibits lack of progress.

%\item Likewise, all other sources of nondeterminism, such as failures and timer expiration, must be modeled using appropriate \psharp APIs. This step allows \psharp to systematically explore all asynchronous event handler interleavings and all other sources of nondeterminism during testing (see \S\ref{sec:psharp:testing}).
\end{enumerate}

@@

\subsection{Specifications}
\label{sec:bg:bugs}

@@

\subsection{Systematic testing}
\label{sec:psharp:testing}

The \psharp runtime is a lightweight layer build on top of the Task Parallel Library (TPL) of .NET that implements the semantics of \psharp: creating state machines, executing them concurrently using the default task scheduler of TPL, and sending events and enqueueing them in the appropriate machines. A key capability of the \psharp runtime is that it can execute in \emph{bug-finding} mode, which systematically tests a \psharp program to find bugs using the safety and liveness monitors that were discussed in \S\ref{sec:bg:bugs}.

@@

The \psharp testing engine will repeatedly execute a program from start to completion, each time exploring a potentially different set of scheduling decisions, until it either reaches a bound (in number of iterations or time), or it hits a property violation. This testing process is fully automatic, has no false-positives (assuming an accurate environment model), and can reproduce found bugs by replaying buggy schedules. After a bug is found, the engine generates a trace that represents the buggy schedule. In contrast to logs typically generated during production, the \psharp trace provides a global order of all communication events and, thus, is easier to debug.

We have implemented two different schedulers inside the \psharp systematic testing engine: a \emph{random} scheduler that makes each nondeterministic choice randomly, and a \emph{probabilistic concurrency testing}~\cite{burckhardt2010pct} (PCT) scheduler. It is straightforward to create a new scheduler by implementing the \texttt{ISchedulingStrategy} interface~\cite{desai2015systematic} exposed by the \psharp libraries. The interface exposes callbacks that are invoked by the \psharp testing engine for taking decisions regarding which machine to schedule next, and can be used for developing both generic and application-specific schedulers.

\textbf{Handling \csharp 5.0 asynchrony}.
Our case studies heavily use the \texttt{async} and \texttt{await} \csharp 5.0 language primitives. An \texttt{async} method is able to perform \texttt{await} on a TPL task \emph{without} blocking the current thread. The \csharp compiler achieves this non-blocking behavior by wrapping the code following an \texttt{await} statement in a TPL \emph{task continuation}, and then returning execution to the caller of the \texttt{async} method. The TPL scheduler executes this task continuation when the task being awaited has completed. In principle, this asynchrony can be modeled using the \psharp primitives for machine creation and message passing. However, in our experience, modeling async/await asynchronous code using \psharp is difficult and time consuming, and thus infeasible for large code bases. To tackle this problem, we developed a solution that automatically captures asynchronously created TPL tasks and wraps them in temporary \psharp machines.

%The \texttt{async} and \texttt{await} \csharp 5.0 primitives were not originally supported in \psharp. To be able to test our case studies, as well as other systems written in \csharp 5.0, we had to extend \psharp with support for intra-machine concurrency (see \S\ref{sec:psharp:testing}).

Our solution involves using a \emph{custom task scheduler}, during systematic testing with \psharp, instead of the default TPL scheduler.
%Assuming that the developer does not explicitly change the task scheduler (something extremely uncommon),
All TPL tasks spawned in the \psharp program will be scheduled for execution in the \psharp task scheduler. These tasks will be intercepted by our custom task scheduler and wrapped inside a special (not available to the programmer) task-machine. When the \psharp testing engine schedules a task-machine for execution, the corresponding task is unwrapped and executed. This approach works nicely with the \texttt{async} and \texttt{await} \csharp 5.0 primitives: when an \texttt{async} method is called, the asynchronous TPL task will be automatically wrapped in a task-machine and scheduled by \psharp; for \texttt{await}, no special treatment is required since the task continuation will also be automatically wrapped and scheduled accordingly.

>>>>

The goal of our work is to detect distributed system bugs \emph{before} they reach production. Distributed systems typically consist of multiple components that interact with each other via \emph{message passing}. If these messages---or unexpected failures and timeouts---are not handled properly, they can lead to subtle bugs. To expose these bugs, we use \psharp~\cite{deligiannis2015psharp}, an extension of the \csharp language that provides: (i) language support for \emph{modeling} the environment of distributed systems written in .NET; and (ii) a \emph{systematic testing engine} that can explore interleavings between \emph{distributed events}, such as the nondeterministic order of message deliveries, failures and timeouts. If the \psharp testing engine identifies a sequence of events that triggers a bug (witnessed by an assertion failure), then exploration terminates, and a trace is reported.

Due to the highly asynchronous nature of distributed storage systems, the number of possible states that these systems can reach is exponentially large. Tools such as \textsc{MoDist}~\cite{yang2009modist} and dBug~\cite{simsa2011dbug} focus on testing unmodified distributed systems, but this can easily lead to state-space explosion when trying to exhaustively explore the entire state-space of a production-scale system, such as the Azure Storage vNext. On the other hand, techniques such as TLA+~\cite{lamport1994temporal} have successfully verified specifications of distributed systems, but they are unable to verify the actual implementation. In this work, we are proposing a solution between the two extremes: systematically test the implementation of a distributed system component against its modeled by \psharp environment. The benefit of our approach is that it can detect bugs in the \emph{actual} implementation, but it can do so in a much reduced state-space. Testing with our approach does not come for free; developers have to invest effort and time into building a test harness and environment using \psharp. However, we argue that developers already have to build test suites and spend significant time testing a distributed system prior to deployment. Using \psharp might require a bit more effort and time, but pays dividends in found bugs (see \S\ref{sec:eval}).

\subsection{A small example}
\label{sec:overview:example}

Draft.

\subsection{Modeling with \psharp}
\label{sec:overview:psharp}

@@

\item The environment of the system must be modeled in \psharp, which offers APIs for obtaining non-deterministically chosen values. These values can be used to simulate failures, timers, etc. (see \S\ref{sec:method:timer} and \S\ref{sec:method:driver}). This step captures external sources of nondeterminism. The \psharp testing engine can then systematically explore all choices, enumerating different asynchronous event interleavings, possible points of failures, timers firing, etc. (see \S\ref{sec:overview:testing}).

\item The criteria for correctness of an execution must be specified. Specifications in \psharp can encode either \emph{safety} or \emph{liveness}~\cite{lamport1977proving} properties (see \S\ref{sec:overview:bugs}). Safety specifications generalize the notion of source code assertions; a safety violation is a finite trace leading to an erroneous state. Liveness specifications generalize nontermination; a liveness violation is an infinite trace that exhibits lack of progress.

%\item Likewise, all other sources of nondeterminism, such as failures and timer expiration, must be modeled using appropriate \psharp APIs. This step allows \psharp to systematically explore all asynchronous event handler interleavings and all other sources of nondeterminism during testing (see \S\ref{sec:overview:testing}).
\end{enumerate}

@@

\subsection{Writing specifications in \psharp}
\label{sec:overview:bugs}

@@

\subsection{Systematic testing using \psharp}
\label{sec:overview:testing}

The \psharp runtime is a lightweight layer build on top of the Task Parallel Library (TPL) of .NET that implements the semantics of \psharp: creating state machines, executing them concurrently using the default task scheduler of TPL, and sending events and enqueueing them in the appropriate machines. A key capability of the \psharp runtime is that it can execute in \emph{bug-finding} mode, which systematically tests a \psharp program to find bugs using the safety and liveness monitors that were discussed in \S\ref{sec:overview:bugs}.

@@

The \psharp testing engine will repeatedly execute a program from start to completion (one execution), each time exploring a potentially different set of scheduling decisions, until it either reaches a bound (in number of executions or time), or it hits a property violation. This testing process is fully automatic, has no false-positives (assuming an accurate environment model), and can reproduce found bugs by replaying buggy schedules. After a bug is found, the engine generates a trace that represents the buggy schedule. In contrast to logs typically generated during production, the \psharp trace provides a global order of all communication events and, thus, is easier to debug.

We have implemented two different schedulers that are responsible for making scheduling and other non-deterministic choices: a \emph{random} scheduler that makes each nondeterministic choice randomly, and a \emph{probabilistic concurrency testing}~\cite{burckhardt2010pct} (PCT) scheduler. It is straightforward to create a new scheduler by implementing the \texttt{ISchedulingStrategy} interface~\cite{desai2015systematic} exposed by the \psharp libraries. The interface exposes callbacks that are invoked by the \psharp testing engine for taking decisions regarding which machine to schedule next, and can be used for developing both generic and application-specific schedulers.

%\textbf{Handling \csharp 5.0 asynchrony}.
%Our case studies heavily use the \texttt{async} and \texttt{await} \csharp 5.0 language primitives. An \texttt{async} method is able to perform \texttt{await} on a TPL task \emph{without} blocking the current thread. The \csharp compiler achieves this non-blocking behavior by wrapping the code following an \texttt{await} statement in a TPL \emph{task continuation}, and then returning execution to the caller of the \texttt{async} method. The TPL scheduler executes this task continuation when the task being awaited has completed. In principle, this asynchrony can be modeled using the \psharp primitives for machine creation and message passing. However, in our experience, modeling async/await asynchronous code using \psharp is difficult and time consuming, and thus infeasible for large code bases. To tackle this problem, we developed a solution that automatically captures asynchronously created TPL tasks and wraps them in temporary \psharp machines.
%
%%The \texttt{async} and \texttt{await} \csharp 5.0 primitives were not originally supported in \psharp. To be able to test our case studies, as well as other systems written in \csharp 5.0, we had to extend \psharp with support for intra-machine concurrency (see \S\ref{sec:overview:testing}).
%
%Our solution involves using a \emph{custom task scheduler}, during systematic testing with \psharp, instead of the default TPL scheduler.
%%Assuming that the developer does not explicitly change the task scheduler (something extremely uncommon),
%All TPL tasks spawned in the \psharp program will be scheduled for execution in the \psharp task scheduler. These tasks will be intercepted by our custom task scheduler and wrapped inside a special (not available to the programmer) task-machine. When the \psharp testing engine schedules a task-machine for execution, the corresponding task is unwrapped and executed. This approach works nicely with the \texttt{async} and \texttt{await} \csharp 5.0 primitives: when an \texttt{async} method is called, the asynchronous TPL task will be automatically wrapped in a task-machine and scheduled by \psharp; for \texttt{await}, no special treatment is required since the task continuation will also be automatically wrapped and scheduled accordingly.

### 2015-12-22 paper.tex
Made some changes mostly relating to the first few comments from reviewer A.
slight add in abstract
update

\usepackage{color}
\definecolor{light-gray}{gray}{0.40}
%\newcommand{\SQComment}[1]{\textcolor{red}{Shaz: #1}}
%\newcommand{\PDComment}[1]{\textcolor{blue}{Pantazis: #1}}
%\newcommand{\MMComment}[1]{\textcolor{cyan}{Matt: #1}}
%\newcommand{\SCComment}[1]{\textcolor{magenta}{Shuo: #1}}
%\newcommand{\CHComment}[1]{\textcolor{blue}{Cheng: #1}}
%\newcommand{\PTComment}[1]{\textcolor{light-gray}{Paul: #1}}

@@

%make title bold and 14 pt font (Latex default is non-bold, 16 pt)
\title{\Large \bf Uncovering Bugs in Distributed Storage Systems\\during Testing (not in Production!)}
%\title{\Large \bf Uncovering Distributed Storage System Bugs during\\Testing (not in Production!)}

@@

\subsection*{Abstract}
Testing distributed storage systems is challenging due to multiple sources of nondeterminism. Conventional testing techniques, such as unit and stress testing, are ineffective in preventing serious but subtle bugs from reaching production. Formal techniques, such as TLA+, can only verify high-level specifications of systems at the level of logic-based models, and fall short of checking the actual executable code. In this paper, we present a new methodology for testing distributed storage systems. Our approach applies advanced systematic testing techniques to thoroughly check that the executable code adheres to its high-level specifications, which significantly improves coverage of important system behaviors.

Our methodology has been applied to three distributed storage systems in the \Microsoft \Azure platform. In the process, numerous bugs were identified, reproduced, confirmed and fixed. These bugs required a subtle combination of concurrency and failures, making them extremely difficult to find with conventional testing techniques. An important advantage of our approach is that a bug is uncovered in a small setting and witnessed by a full system trace, which dramatically increases the productivity of debugging.

@@

\section{Our Approach to Testing Distributed Systems}
\label{sec:overview}

>>>>

\usepackage{color}
\definecolor{light-gray}{gray}{0.40}
\definecolor{orange}{rgb}{1,0.5,0}
%\newcommand{\SQComment}[1]{\textcolor{red}{Shaz: #1}}
%\newcommand{\PDComment}[1]{\textcolor{blue}{Pantazis: #1}}
%\newcommand{\MMComment}[1]{\textcolor{cyan}{Matt: #1}}
%\newcommand{\SCComment}[1]{\textcolor{magenta}{Shuo: #1}}
%\newcommand{\CHComment}[1]{\textcolor{blue}{Cheng: #1}}
\newcommand{\PTComment}[1]{\textcolor{orange}{Paul: #1}}

@@

%make title bold and 14 pt font (Latex default is non-bold, 16 pt)
\title{\Large \bf Uncovering Bugs in Distributed Storage Systems\\during Testing (not in Production!)}

@@

\subsection*{Abstract}
Testing distributed systems is challenging due to multiple sources of nondeterminism. Conventional testing techniques, such as unit, integration and stress testing, are ineffective in preventing serious but subtle bugs from reaching production. Formal techniques, such as TLA+, can only verify high-level specifications of systems at the level of logic-based models, and fall short of checking the actual executable code. In this paper, we present a new methodology for testing distributed systems. Our approach applies advanced systematic testing techniques to thoroughly check that the executable code adheres to its high-level specifications, which significantly improves coverage of important system behaviors.

Our methodology has been applied to three distributed storage systems in the \Microsoft \Azure cloud computing platform. In the process, numerous bugs were identified, reproduced, confirmed and fixed. These bugs required a subtle combination of concurrency and failures, making them extremely difficult to find with conventional testing techniques. An important advantage of our approach is that a bug is uncovered in a small setting and witnessed by a full system trace, which dramatically increases the productivity of debugging.

@@

\section{Overview of our Approach}
%\section{Our Approach to Testing Distributed Systems}
\label{sec:overview}
